---
title: "What drives the functional response to disturbance in Paracou?"
output: 
  html_document:
    theme: yeti
    keep_md: true
---

```{r global_options, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning=FALSE, message=FALSE, results='hide',
                      root.dir = "C:/Users/camille.piponiot/gitR/functional_trajectories_paracou")
library(rstan)
rstan_options(auto_write = TRUE)
options(mc.cores = parallel::detectCores())

library(data.table)
library(ggplot2)
library(gridExtra)
library(tidyr)

inference_stan = FALSE
```

# Data visualisation

```{r, traits_traj, fig.height=10, fig.width=10}
# to get the data: source("C:/Users/camille.piponiot/Google Drive/biodiversity/codes/get_metadata_paracou.R")

load("data/genusData.Rdata")
load("data/plotInfo.Rdata")
name_traits = c("woodDensity","logSeedMass","SLA","DBH95")


### reorganize data -> gather dataframes
genusData_melt = melt(genusData, variable.name = "trait", value.name = "value", 
                      measure.vars = grep("D|Seed|SLA",colnames(genusData)), 
                      variable.factor = FALSE)
dfTraits = genusData_melt[,.(WMT = weighted.mean(value, agb)),.(idplot, year, trait, cohort)]

## add plot info (plot size, treatment)
dfTraits = merge(dfTraits, plotInfo)

## initial value
md_0 = dfTraits[year < 1987,.(WMT0 = mean(WMT)),.(idplot, trait, treat)]

```

## Traits trajectory 

Traits chosen: 

- $DBH95$: DBH 95th percentile (per species) as a proxy of the stature

- $logSeedMass$: median seed mass (log-transformed) as a proxy of the dispersal strategy

- $SLA$: median specific leaf area, as a proxy of the assimilation strategy and leaf economic spectrum

- $WD$: median wood density, as a proxy of growth rate and mechanical support

Cohorts: 

- Survivors, ie trees that survived logging

- Recruits, ie trees that passed the 10 cm DBH threshold after logging

Here are represented for each cohort the mean trait value, weighted by the estimated above ground biomass (AGB) of each individual.

```{r recr_traj,fig.height=10, fig.width=10}

dfTraits$cohort <- factor(dfTraits$cohort)
levels(dfTraits$cohort) <- c("Recruits", "Survivors")

ggplot(dfTraits, aes(x = year - 1986, y = WMT, group = idplot, colour = treat)) + 
  geom_line() + geom_point() + 
  geom_point(data = md_0, aes(x = 0, y = WMT0), pch = 1, cex = 5) + 
  ylab("MWT") + xlab("Time since logging (yr)") +
  facet_grid(trait~cohort, scales = "free")

```

For $D$: there seems to be no effect of logging gaps on recruitment, the only effect is on survivors (direct removal of big trees + higher growth rates of big survivors): worth including in the study? 

We can exclude $D$ from the study later. 

## Recruits most common genuses and their impact on mean weighted traits {.tabset .tabset-pills}

### AGB per genus

```{r, fig.height=10, fig.width=12}

recrData = merge(genusData, genusData[cohort=="recr",.(agbR = sum(agb)),.(idplot, year, cohort)])

## genuses that at least once represent 10% of the total biomass of recruits
genus_common = unique(as.character(subset(recrData, agb/agbR > 0.15 & !(genus %in% c("Oenocarpus", "Indet")))$genus)) 

recrData$genus2 = as.character(recrData$genus)
recrData$genus2[!(recrData$genus %in% genus_common)] <- "other"

recrData_agb = recrData[, .(agb = sum(agb)), .(genus2,idplot,year,agbR)]
recrData = merge(recrData_agb, recrData[, .(woodDensity = weighted.mean(woodDensity, agb),
                                        logSeedMass = weighted.mean(logSeedMass, agb),
                                        SLA = weighted.mean(SLA, agb)), .(genus2)])

recrData = merge(recrData, plotInfo, by="idplot")
recrData = merge(recrData, data.table(treat = c("ctrl","CL","silv","silv+"), treatment = paste("T",0:3,sep="")), by="treat")
recrData$bloc = rep(1:3,each=4)[as.numeric(recrData$idplot)]

recrData$genus2 = factor(recrData$genus2, levels = c(as.character(genus_common), "other") )
ggplot(recrData, aes(x=year, y=agb, fill=genus2)) + geom_area(colour = 1) + facet_grid(treatment ~ bloc)

## mean traits (weighted by biomass) of recruited genuses in control plots -> reference (= white color in plots)
ref = recrData[treat=="ctrl", .(woodDensity = weighted.mean(woodDensity, agb),
                                logSeedMass = weighted.mean(logSeedMass, agb),
                                SLA = weighted.mean(SLA, agb))]
```

### Proportion of total AGB 

```{r, fig.height=10, fig.width=12}
ggplot(recrData, aes(x = year, y = agb/agbR, fill = genus2)) + geom_area(colour = 1) + facet_grid(treatment ~ bloc, scales= "free")
```

### Seed mass

```{r, fig.height=10, fig.width=12}
ggplot(recrData, aes(x = year, y = agb/agbR, group = genus2, fill = logSeedMass)) + geom_area(colour = 1) + facet_grid(treatment ~ bloc, scales= "free") + scale_fill_gradient2(low = "firebrick4", mid = "white", high = "#104E8B", midpoint = ref$logSeedMass)
```

### SLA

```{r, fig.height=10, fig.width=12}
ggplot(recrData, aes(x = year, y = agb/agbR, group = genus2, fill = SLA)) + geom_area(colour = 1) + facet_grid(treatment ~ bloc, scales= "free") + scale_fill_gradient2(low = "firebrick4", mid = "white", high = "#104E8B", midpoint = ref$SLA)
```

### WD

```{r, fig.height=10, fig.width=12}
ggplot(recrData, aes(x = year, y = agb/agbR, group = genus2, fill = woodDensity)) + geom_area(colour = 1) + facet_grid(treatment ~ bloc, scales= "free") + scale_fill_gradient2(low = "firebrick4", mid = "white", high = "#104E8B", midpoint = ref$woodDensity)
```

### Conclusions 

Early pioneer = cecropias (SLA 0 // seed mass -- // WD -- )

Late pioneer = ingas + miconias + tapiriras + vismias (SLA ++ // seed mass - // WD - )

- seed mass follows exactly cecropias dynamics: peaks quickly (low $tmax$, around 10 years) and recovers fast (high $\theta$)

- WD peaks quickly (low $tmax$) but stays low for long (low $\theta$)

- SLA peaks after a long time (high $tmax$) but stays low for longer (low $\theta$)

## Identity of trees logged and devitalised

```{r mortSurvTraits, fig.height = 3, fig.width = 10}
load("data/mortLogging.Rdata")

dfMort = melt(mortLogging, id.vars = c("eventMort","idplot","agb"), variable.name = "trait", value.name = "WMT")

dfMort = merge(dfMort, md_0, by = c("idplot", "trait"))

ggplot(dfMort, aes(y = (WMT - WMT0)/WMT0*100, x = eventMort, colour=treat )) + 
  geom_hline(yintercept = 0, lty = 2) + 
  geom_boxplot() + 
  ylab("Difference between killed trees'\nand prelogging trait values (%)") +
  xlab("Type of tree mortality") + 
  facet_wrap( ~ trait, scales = "free")
```
