---
title: "What drives the functional response to logging in Amazonia?"
output: 
  html_document:
    theme: yeti
    keep_md: true
    toc: true
    toc_float: true
---

```{r global_options, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning=FALSE, message=FALSE, results='hide',
                      root.dir = "C:/Users/camille.piponiot/gitR/functional_trajectories_TmFO")
library(rstan)
rstan_options(auto_write = TRUE)
options(mc.cores = parallel::detectCores())

library(data.table)
library(ggplot2)
library(gridExtra)
library(tidyr)
library(knitr)
library(ggpubr)

inference_stan = FALSE
```


In this study we model the functional resistance and resilience of trees recruitement to disturbance (here selective logging), and their spatial variation in Amazonia. 

# Data visualisation

## by cohort
```{r, traits_traj, fig.height=10, fig.width=10}
# to get the data: source("C:/Users/camille.piponiot/Google Drive/biodiversity/codes/get_metadata.R")
load("data/sizeData_TmFO.Rdata")
load("data/plotInfo_TmFO.Rdata")
name_traits = c("woodDensity","logSeedMass","SLA","DBH95")

plotInfo = merge(plotInfo, sizeData[,.(tlog = min(year)), .(idplot)], by = "idplot")
plotInfo$tlog[plotInfo$site == "prc"] = 1986
plotInfo$tlog[plotInfo$site == "tor"] = 2006
plotInfo$tlog[plotInfo$site == "cum"] = 2004

### reorganize data -> gather dataframes
sizeData_melt = melt(sizeData, variable.name = "trait", value.name = "value", 
                     measure.vars = grep("D|Seed|SLA",colnames(sizeData)), 
                     variable.factor = FALSE)
dfTraits = sizeData_melt[,.(WMT = weighted.mean(value, agb), agb_tot = sum(agb)),.(idplot, year, trait, size)]

## add plot info (plot size, treatment)
dfTraits = merge(dfTraits, plotInfo, by = c("idplot"))
dfTraits[, agb := agb_tot/idplot.size]

## initial value
md_0 = dfTraits[year <= tlog,.(WMT0 = mean(WMT)),.(site, idplot, trait, treat)]

```

## Traits trajectory {.tabset .tabset-fade .tabset-pills}

Traits chosen: 

- $DBH95$: DBH 95th percentile (per species) as a proxy of the stature

- $logSeedMass$: median seed mass (log-transformed) as a proxy of the dispersal strategy

- $SLA$: median specific leaf area, as a proxy of the assimilation strategy and leaf economic spectrum

- $WD$: median wood density, as a proxy of growth rate and mechanical support

We separate two size classes: 

- Small trees, ie trees $\leq$ 35 cm dbh

- Big trees, ie trees $>$ 35 cm dbh

Here are represented for each size class the mean trait value, weighted by the estimated above ground biomass (AGB) of each individual.

```{r}
dfTraits$size <- factor(dfTraits$size)
levels(dfTraits$size) <- c("small", "big")

graph_traits <- function(sit) {
  ggplot(subset(dfTraits, site == sit), aes(x = year - tlog, y = WMT, 
                                            group = idplot, colour = treat)) + 
    geom_line() + geom_point() + 
    geom_point(data = subset(md_0, site == sit), aes(x = 0, y = WMT0), pch = 1, cex = 5) + 
    ylab("MWT") + xlab("Time since logging (yr)") +
    facet_grid(trait ~ size, scales = "free")
}

tmfo_sites <- as.character(unique(dfTraits$site))
```

### Jenaro 
```{r, fig.height=8, fig.width = 10}
graph_traits("bsl")
```

### Chico Bocao 
```{r, fig.height=10, fig.width = 10}
graph_traits("chb")
```

### Cumaru 
```{r, fig.height=10, fig.width = 10}
graph_traits("cum")
```

### INPA 
```{r, fig.height=10, fig.width = 10}
graph_traits("inp")
```

### Iracema 
```{r, fig.height=10, fig.width = 10}
graph_traits("ira")
```

### La Chonta 
```{r, fig.height=10, fig.width = 10}
graph_traits("lch")
```

### Peteco 
```{r, fig.height=10, fig.width = 10}
graph_traits("pet")
```

### Paracou 
```{r, fig.height=10, fig.width = 10}
graph_traits("prc")
```

### Paragominas 
```{r, fig.height=10, fig.width = 10}
graph_traits("prg")
```

### Tabocal 
```{r, fig.height=10, fig.width = 10}
graph_traits("tbc")
```

### Tapajos 
```{r, fig.height=10, fig.width = 10}
graph_traits("tpj")
```

### Tortue 
```{r, fig.height=10, fig.width = 10}
graph_traits("tor")
```


## Recruits most common genuses and their impact on mean weighted traits {.tabset .tabset-pills}

```{r, fig.height=20, fig.width=12}
load("data/genusData_TmFo.Rdata")
recrData = merge(genusData, genusData[cohort=="recr",.(agbR = sum(agb)),
                                      .(idplot, year, cohort)])

recrData = merge(recrData, plotInfo, by=c("site", "idplot"))
```

```{r}
graph_genus <- function(sit) {
  
  dat = subset(recrData, site == sit)
  ## 5 most common genuses
  genus_common = dat[ !(genus %in% c("Oenocarpus", "Indet", NA,"ni")), .(agb = sum(agb)),.(genus)][order(agb, decreasing = TRUE)][1:5]$genus
  dat$genus2 = as.character(dat$genus)
  dat$genus2[!(dat$genus %in% genus_common)] <- "other"
  dat$genus2 = factor(dat$genus2, levels = c(as.character(genus_common), "other") )
  
  dat_agb = dat[, .(agb = sum(agb)), .(genus2,idplot,year,agbR)]
  dat = merge(dat_agb, dat[, .(woodDensity = weighted.mean(woodDensity, agb),
                               logSeedMass = weighted.mean(logSeedMass, agb),
                               SLA = weighted.mean(SLA, agb)), .(genus2)])
  
  g1 = ggplot(dat, aes(x=year, y=agb, fill=genus2)) + 
    geom_area(colour = 1) + facet_wrap( ~ idplot)
  g1
  ## mean traits (weighted by biomass) of recruited genuses in control plots -> reference (= white color in plots)
  # ref = dat[treat=="ctrl", .(woodDensity = weighted.mean(woodDensity, agb),
  #                            logSeedMass = weighted.mean(logSeedMass, agb),
  #                            SLA = weighted.mean(SLA, agb))]
}

```

genus2: genuses with > 10% of the recruits biomass

### Jenaro 
```{r, fig.height=4, fig.width = 10}
graph_genus("bsl")
```

### Chico Bocao 
```{r, fig.height=4, fig.width = 6}
graph_genus("chb")
```

### Cumaru 
```{r, fig.height=4, fig.width = 6}
graph_genus("cum")
```

### INPA 
```{r, fig.height=10, fig.width = 10}
graph_genus("inp")
```

### Iracema 
```{r, fig.height=4, fig.width = 6}
graph_genus("ira")
```

### La Chonta 
```{r, fig.height=8, fig.width = 10}
graph_genus("lch")
```

### Peteco 
```{r, fig.height=4, fig.width = 6}
graph_genus("pet")
```

### Paracou 
```{r, fig.height=8, fig.width = 10}
graph_genus("prc")
```

### Paragominas 
```{r, fig.height=4, fig.width = 10}
graph_genus("prg")
```

### Tabocal 
```{r, fig.height=4, fig.width = 6}
graph_genus("tbc")
```

### Tapajos 
```{r, fig.height=4, fig.width = 10}
graph_genus("tpj")
```

### Tortue 
```{r, fig.height=4, fig.width = 10}
graph_genus("tor")
```


## Calibrating the stan model

For each trait $k$, the mean biomass-weighted trait of small trees (MWT) trajectory at census $c$ in plot $p$ in site $s$ was modelled as:

$$  dT_{k,c,p,s} = \frac{T_{k,c,p,s} - T0_{k,p,s}}{T0_{k,p,s}} \sim \mathcal{N}\left( \mu_{k,c,p,s}\text{ , } \left(\frac{\sigma_k}{size_p}\right)^2\right)$$

with 

$$ \mu_{k,c,p,s} = \left(\Delta_{k,p,s} \cdot  \frac{t_c}{tmax_{k,p,s}} \cdot exp\left(1-\frac{t_c}{tmax_{k,p,s}}\right)\right)^{\theta_{k,p,s}} $$

- $c$ the census, and $t_c$ the time since logging (in years) at census $c$, 

- $T0_{k,p,s}$ the pre-logging trait value of the community in plot $p$ in site $s$.

- $\Delta_{k,p,s}$ is the maximum value (relative to the final value) of the hump that follows logging and the creation of logging gaps. $\Delta_{k,p,s} \sim \mathcal{N} (\mu_{\Delta ks}, \sigma_{\Delta}^2)$, with $\mu_{\Delta ks}$ and $\sigma_{\Delta}$ are respectively the mean and standard deviation of $\Delta$ for trait $k$ in site $s$.

- $tmax_{k,p,s}$ is the time when the maximum value of the hump is reached. $tmax_{k,p} = 0$ for control plots and in logged plots $tmax_{k,p} \sim \mathcal{N} (\mu m_{ k}, \sigma m^2)$ where $\mu m_{k}$ and $\sigma m$ are respectively the mean and standard deviation of $tmax$ for trait $k$.

- $\theta_k$ is a shape parameter that controls the width of the hump; when it increases, the hump is narrower.  

- $size_p$ is the size of plot $p$

```{r}

if (inference_stan){
  
  df = subset(dfTraits, size=="small" & treat != "ctrl")
  name_traits = unique(df$trait)
  df = dcast(df, site + idplot + tlog + year + idplot.size ~ trait , value.var = "WMT")
  df0 = subset(df, year <= tlog)
  df = subset(df, year > tlog)
  
  N = nrow(df)
  K = length(name_traits)
  np = as.numeric(as.factor(as.character(df$idplot)))
  P = length(unique(np))
  ns = as.numeric(as.factor(as.character(df$site)))
  S = length(unique(ns))
  ps = unique(cbind(ns, np))[,1]
  plot_size = unique(cbind(df$idplot.size, np))[,1]
  
  t = df$year - df$tlog
  TR = as.matrix(df[, name_traits, with = FALSE])
  dfT0 = melt(df0, measure.vars = name_traits, id.vars = "idplot")
  dfT0 = dfT0[,.(value = mean(value)),.(variable,idplot)]
  T0 = as.matrix(dcast(dfT0, idplot ~variable, value.var = "value")[,-"idplot"])
  
  small_model = stan("codes/recruits_model.stan")
  traceplot(small_model, "lp__")
  pars_small = data.table(do.call(cbind, rstan::extract(small_model, 
                                                        pars = c("sigma_Delta", "sigma_tmax", 
                                                                 "mu_theta", "sigma_theta", "lp__" )))[1:1000])
  pars_small$iter = 1:1000
  pars_smallb = data.table(Delta = c(rstan::extract(small_model, pars= "Delta")[[1]][1:1000,,]), 
                           theta = c(rstan::extract(small_model, pars= "theta")[[1]][1:1000,,]),
                           tmax = c(rstan::extract(small_model, pars= "tmax")[[1]][1:1000,,]), 
                           iter = rep(1:1000, P*K), 
                           p = rep(rep(1:P, each = 1000), K), 
                           s = rep(rep(ps, each = 1000), K), 
                           k = rep(1:K, each = 1000*P))
  pars_small = merge(pars_small, pars_smallb, by = "iter")
  pars_smallb = data.table(sigmaT = c(rstan::extract(small_model, pars= "sigmaT")[[1]][1:1000,]),
                           mu_Delta = c(rstan::extract(small_model, pars= "mu_Delta")[[1]][1:1000,]),
                           iter = rep(1:1000, K), 
                           k = rep(1:K, each = 1000))
  pars_small = merge(pars_small, pars_smallb, by = c("iter", "k"))
  pars_smallb = data.table(mu_tmax = c(rstan::extract(small_model, pars= "mu_tmax")[[1]][1:1000,,]), 
                           iter = rep(1:1000, S*K), 
                           s = rep(rep(1:S, each = 1000), K), 
                           k = rep(1:K, each = 1000*S))
  pars_small = merge(pars_small, pars_smallb, by = c("iter","s","k"))
  pars_small$trait = name_traits[pars_small$k]
  pars_small$site = levels(as.factor(as.character(df$site)))[pars_small$s]
  pars_small$idplot = levels(as.factor(as.character(df$idplot)))[pars_small$p]
  save(pars_small, file = "data/parsSmall.Rdata")
} else { load("data/parsSmall.Rdata")}

```

## Parameters {.tabset .tabset-pills}
(check parameters correlation)

### Delta
```{r, fig.height = 6, fig.width = 12}
ggplot(pars_small, aes(x = idplot, y = Delta)) + geom_hline(yintercept = 0, lty=2) +
  geom_boxplot() + facet_grid(trait ~ site, scales = "free")
```

### tmax
```{r, fig.height = 6, fig.width = 12}
ggplot(pars_small, aes(x = idplot, y = tmax)) +
  geom_boxplot() + facet_grid(trait ~ site, scales = "free")
```

### theta
```{r, fig.height = 6, fig.width = 12}
ggplot(pars_small, aes(x = idplot, y = theta)) +
  geom_boxplot() + facet_grid(trait ~ site, scales = "free")
```

### theta vs tmax
```{r, fig.height = 6, fig.width = 12}
ggplot(pars_small, aes(x = tmax, y = theta, col = idplot)) + 
  geom_point(alpha = 0.2) + facet_grid(trait ~ site, scales = "free") + theme(legend.position = "none")
```

### Delta vs tmax
```{r, fig.height = 6, fig.width = 12}
ggplot(pars_small, aes(x = tmax, y = Delta, col = idplot)) + 
  geom_point(alpha = 0.2) + facet_grid(trait ~ site, scales = "free") + theme(legend.position = "none")
```

### theta vs Delta
```{r, fig.height = 6, fig.width = 12}
ggplot(pars_small, aes(x = Delta, y = theta, col = idplot)) + 
  geom_point(alpha = 0.2) + facet_grid(trait ~ site, scales = "free") + theme(legend.position = "none")
```

## Predictions {.tabset .tabset-pills}

```{r}
df_pred = data.table(expand.grid(iter = 1:100, trait = unique(dfTraits$trait), t = 0:35, idplot = unique(dfTraits$idplot)))
df_pred = merge(df_pred, pars_small[,c("iter", "idplot", "trait","site", "Delta", "tmax", "theta")], by = c("trait","iter","idplot"))
dfT0 = subset(dfTraits, size=="small" & year <= tlog)[,.(T0 = mean(WMT)), .(idplot, trait)]
df_pred = merge(df_pred, dfT0, by = c("idplot","trait"))

df_pred[, dtr := Delta * ( t / tmax * exp( 1 - t / tmax))^theta ]
df_ci = df_pred[,.(inf = quantile(dtr, 0.025), 
                   med = quantile(dtr, 0.5), 
                   sup = quantile(dtr, 0.975)), .(trait,idplot,t)]
df_ci = merge(df_ci, plotInfo, by = "idplot")

dfTraits = merge(dfTraits, dfT0, by = c("idplot","trait"))
dfTraits[, dtr := (WMT-T0)/T0]
```

```{r}
graph_pred = function(sit){
  ggplot(subset(dfTraits, size=="small" & site == sit & treat!="ctrl"), aes(x = year - tlog, y = dtr, colour = idplot)) + 
    geom_ribbon(data = subset(df_ci, site == sit), aes(x = t, ymin = inf, ymax = sup, fill = idplot, y = med), colour = NA, alpha = 0.2) +
    geom_line(data = subset(df_ci, site == sit), aes(x = t, y = med, colour = idplot), lty = 2) +
    geom_point() + facet_grid(trait ~ treat, scales = "free") +
    labs(y = "Relative trait change", x = "Time since logging (yr)") +
    theme(legend.position = "none")
}
```

### Jenaro 
```{r, fig.height=8, fig.width = 4}
graph_pred("bsl")
```

### Chico Bocao 
```{r, fig.height=8, fig.width = 4}
graph_pred("chb")
```

### Cumaru 
```{r, fig.height=8, fig.width = 4}
graph_pred("cum")
```

### INPA 
```{r, fig.height=8, fig.width = 10}
graph_pred("inp")
```

### Iracema 
```{r, fig.height=8, fig.width = 4}
graph_pred("ira")
```

### La Chonta 
```{r, fig.height=8, fig.width = 10}
graph_pred("lch")
```

### Peteco 
```{r, fig.height=8, fig.width = 4}
graph_pred("pet")
```

### Paracou 
```{r, fig.height=8, fig.width = 10}
graph_pred("prc")
```

### Paragominas 
```{r, fig.height=8, fig.width = 10}
graph_pred("prg")
```

### Tabocal 
```{r, fig.height=8, fig.width = 4}
graph_pred("tbc")
```

### Tapajos 
```{r, fig.height=8, fig.width = 8}
graph_pred("tpj")
```

### Tortue 
```{r, fig.height=8, fig.width = 8}
graph_pred("tor")
```


## Effect of covariates {.tabset .tabset-pills}


```{r}
# source("codes/get_env_covariates.R")
load("data/site_coord.Rdata")

pars_melt = melt(pars_small, measure.vars = c("Delta","tmax","theta"), 
                 id.vars = c("iter","site","trait","idplot"), variable.name = "parameter")
df_covar = merge(pars_melt[,.(mean = mean(value), sd = sd(value)), .(site, trait, parameter, idplot)],
                 site_coord, by = "site")

## agb loss
df_agb = dfTraits[trait == "SLA" & treat != "ctrl",
                  .(agb = sum(agb_tot)), .(site,idplot,year,tlog)]
df_dist = df_agb[,.(agb0 = mean(agb[year <= tlog]), 
                    agbmin = min(agb[year>tlog & year <= tlog + 4])),.(idplot)]
df_dist[, loss_rel := (agb0-agbmin)/agb0]
df_dist[, loss_abs := (agb0-agbmin)]
df_covar = merge(df_covar,  df_dist, by = "idplot")

df_covar = melt(df_covar, id.vars = c("site", "idplot","trait","parameter","mean","sd"), 
                measure.vars = c(colnames(site_coord)[6:length(site_coord)],"loss_rel","loss_abs"), 
                variable.name = "covariate")

df_covar$covar_typ = "soil"
df_covar$covar_typ[df_covar$covariate %in% c("prec","seas","dry","rad")] = "clim"
df_covar$covar_typ[grep("loss",df_covar$covariate)] = "dist"

```

### Disturbance {.tabset .tabset-pills}

#### Delta
```{r, fig.height=8, fig.width=6}
ggplot(subset(df_covar, parameter == "Delta" & covar_typ == "dist"), 
       aes(x = value, y = mean, size = 1/sd, weight = 1/sd, colour=site)) + 
  geom_smooth(method = "lm", formula = y ~ x-1) + 
  geom_point() + facet_grid( trait ~ covariate, scale = "free") + theme(legend.position = "top") 
```


#### abs(Delta)
```{r, fig.height=8, fig.width=6}
ggplot(subset(df_covar, parameter == "Delta" & covar_typ == "dist"), aes(x = value, y = abs(mean), size = 1/sd, weight = 1/sd)) + 
  geom_smooth(method = "lm", formula = y ~ poly(x, 2)) + 
  geom_point(aes(colour = site)) + facet_grid( trait ~ covariate, scale = "free") + theme(legend.position = "top") 
```


#### tmax
```{r, fig.height=8, fig.width=6}
ggplot(subset(df_covar, parameter == "tmax" & covar_typ == "dist"), aes(x = value, y = mean, size = 1/sd, weight = 1/sd)) + 
  geom_smooth(method = "lm", formula = y ~ poly(x, 2)) + 
  geom_point(aes(colour = site)) + facet_grid( trait ~ covariate, scale = "free") + theme(legend.position = "top") 
```

### Climate {.tabset .tabset-pills}

#### Delta
```{r, fig.height=8, fig.width=10}
ggplot(subset(df_covar, parameter == "Delta" & covar_typ == "clim"), aes(x = value, y = mean, size = 1/sd, weight = 1/sd)) + 
  geom_smooth(method = "lm", formula = y ~ poly(x, 2)) + 
  geom_point(aes(colour = site)) + facet_grid( trait ~ covariate, scale = "free") + theme(legend.position = "top") 
```


#### abs(Delta)
```{r, fig.height=8, fig.width=10}
ggplot(subset(df_covar, parameter == "Delta" & covar_typ == "clim"), aes(x = value, y = abs(mean), size = 1/sd, weight = 1/sd)) + 
  geom_smooth(method = "lm", formula = y ~ poly(x, 2)) + 
  geom_point(aes(colour = site)) + facet_grid( trait ~ covariate, scale = "free") + theme(legend.position = "top") 
```


#### tmax
```{r, fig.height=8, fig.width=10}
ggplot(subset(df_covar, parameter == "tmax" & covar_typ == "clim"), aes(x = value, y = mean, size = 1/sd, weight = 1/sd)) + 
  geom_smooth(method = "lm", formula = y ~ poly(x, 2)) + 
  geom_point(aes(colour = site)) + facet_grid( trait ~ covariate, scale = "free") + theme(legend.position = "top") 
```

### Soil {.tabset .tabset-pills}

#### Delta
```{r, fig.height=8, fig.width=10}
ggplot(subset(df_covar, parameter == "Delta" & covar_typ == "soil"), aes(x = value, y = mean, size = 1/sd, weight = 1/sd)) + 
  geom_smooth(method = "lm", formula = y ~ poly(x, 2)) + 
  geom_point(aes(colour = site)) + facet_grid( trait ~ covariate, scale = "free") + theme(legend.position = "top") 
```


#### abs(Delta)
```{r, fig.height=8, fig.width=10}
ggplot(subset(df_covar, parameter == "Delta" & covar_typ == "soil"), aes(x = value, y = abs(mean), size = 1/sd, weight = 1/sd)) + 
  geom_smooth(method = "lm", formula = y ~ poly(x, 2)) + 
  geom_point(aes(colour = site)) + facet_grid( trait ~ covariate, scale = "free") + theme(legend.position = "top") 
```


#### tmax
```{r, fig.height=8, fig.width=10}
ggplot(subset(df_covar, parameter == "tmax" & covar_typ == "soil"), aes(x = value, y = mean, size = 1/sd, weight = 1/sd)) + 
  geom_smooth(method = "lm", formula = y ~ poly(x, 2)) + 
  geom_point(aes(colour = site)) + facet_grid( trait ~ covariate, scale = "free") + theme(legend.position = "top") 
```


> comparer distribution de tmax et Delta **entre** traits

> Lien Delta et intensité de perturbation <=> resilience

> lien est-il spatialement défini? 

## Environmental covariates correlation {.tabset .tabset-pills}

```{r}
library(GGally)
ggcorr(site_coord[,6:14], nbreaks = 4, label = TRUE)

# library(ggbiplot)
rownames(site_coord) = site_coord$site
biplot(princomp(site_coord[,6:9], cor=TRUE))
biplot(princomp(site_coord[,10:14], cor=TRUE))
```

### Option 1
```{r}
ggcorr(site_coord[,c(6,7,13,14)], nbreaks = 4, label = TRUE)
```

### Option 2
```{r}
ggcorr(site_coord[,c(8,9,11,13:14)], nbreaks = 4, label = TRUE)
```

### Maps
```{r}
site_coord_melt = melt(site_coord, id.vars = c("site","Long","Lat"), 
                       measure.vars = colnames(site_coord)[6:14], 
                       variable.name = "covariate")
library(plyr)
col <- c('yellow', 'orange', 'red')
d_ply(site_coord_melt, .(covariate), function(cv){
  g = ggplot(cv , aes(x = Long, y = Lat, colour = value)) + borders() + 
  geom_point(size = 10) + lims(x=c(-80,-40), y =c(-20,10))  + coord_equal() +
    labs(colour=unique(cv$covariate)) + scale_colour_gradientn(colours = col)
  print(g)
})
```


## Variable selection - Delta {.tabset .tabset-pills}

```{r, fig.height=10, fig.width=10}

df_Delta = subset(df_covar, parameter == "Delta")
df_Delta = dcast(unique(df_Delta[,c("site","idplot","mean","sd","covariate","value","trait")]), 
                 site + idplot + trait + mean + sd ~ covariate, value.var = "value")
cov_sc = data.table(scale(site_coord[,6:14]))
colnames(cov_sc) = paste0(colnames(cov_sc), "_sc")
cov_sc$site = rownames(site_coord)
df_Delta = merge(df_Delta, cov_sc, by = "site")

library(sjPlot)
library(RcmdrMisc)
```

### Option 1 {.tabset .tabset-pills}

#### DBH95
```{r}
reg = lm(formula = mean ~ (prec_sc + seas_sc + Depth_sc + CEC_sc) : loss_rel + loss_rel - 1, 
          weights = 1/sd, data = subset(df_Delta, trait == "DBH95"))
select = stepwise(reg, criterion = "BIC")
```

```{r, results = "asis"}
tab_model(select)
```

#### SLA
```{r}
reg = lm(formula = mean ~ (prec_sc + seas_sc + Depth_sc + CEC_sc) : loss_rel + loss_rel - 1, 
          weights = 1/sd, data = subset(df_Delta, trait == "SLA"))
select = stepwise(reg, criterion = "BIC")
```

```{r, results = "asis"}
tab_model(select)
```

#### logSeedMass
```{r}
reg = lm(formula = mean ~ (prec_sc + seas_sc + Depth_sc + CEC_sc) : loss_rel + loss_rel - 1, 
          weights = 1/sd, data = subset(df_Delta, trait == "logSeedMass"))
select = stepwise(reg, criterion = "BIC")
```

```{r, results = "asis"}
tab_model(select)
```

#### woodDensity
```{r}
reg = lm(formula = mean ~ (prec_sc + seas_sc + Depth_sc + CEC_sc) : loss_rel + loss_rel - 1, 
          weights = 1/sd, data = subset(df_Delta, trait == "woodDensity"))
select = stepwise(reg, criterion = "BIC")
```

```{r, results = "asis"}
tab_model(select)
```

### All covariates {.tabset .tabset-pills}

#### DBH95
```{r}
reg = lm(formula = mean ~ (prec_sc + seas_sc + dry_sc + rad_sc + Depth_sc + Psand_sc + CEC_sc + BkD_sc + CFr_sc) : loss_rel + loss_rel - 1, 
          weights = 1/sd, data = subset(df_Delta, trait == "DBH95"))
select = stepwise(reg, criterion = "BIC")
```

```{r, results = "asis"}
tab_model(select)
```

#### SLA
```{r}
reg = lm(formula = mean ~ (prec_sc + seas_sc + dry_sc + rad_sc + Depth_sc + Psand_sc + CEC_sc + BkD_sc + CFr_sc) : loss_rel + loss_rel - 1, 
          weights = 1/sd, data = subset(df_Delta, trait == "SLA"))
select = stepwise(reg, criterion = "BIC")
```

```{r, results = "asis"}
tab_model(select)
```

#### logSeedMass
```{r}
reg = lm(formula = mean ~ (prec_sc + seas_sc + dry_sc + rad_sc + Depth_sc + Psand_sc + CEC_sc + BkD_sc + CFr_sc) : loss_rel + loss_rel - 1, 
          weights = 1/sd, data = subset(df_Delta, trait == "logSeedMass"))
select = stepwise(reg, criterion = "BIC")
```

```{r, results = "asis"}
tab_model(select)
```

#### woodDensity
```{r}
reg = lm(formula = mean ~ (prec_sc + seas_sc + dry_sc + rad_sc + Depth_sc + Psand_sc + CEC_sc + BkD_sc + CFr_sc) : loss_rel + loss_rel - 1, 
          weights = 1/sd, data = subset(df_Delta, trait == "woodDensity"))
select = stepwise(reg, criterion = "BIC")
```

```{r, results = "asis"}
tab_model(select)
```

### Option 2 {.tabset .tabset-pills}

#### DBH95
```{r}
reg = lm(formula = mean ~ (dry_sc + rad_sc + Depth_sc + Psand_sc + CEC_sc) : loss_rel + loss_rel - 1, 
          weights = 1/sd, data = subset(df_Delta, trait == "DBH95"))
select = stepwise(reg, criterion = "BIC")
```

```{r, results = "asis"}
tab_model(select)
```

#### SLA
```{r}
reg = lm(formula = mean ~ (dry_sc + rad_sc + Depth_sc + Psand_sc + CEC_sc) : loss_rel + loss_rel - 1, 
          weights = 1/sd, data = subset(df_Delta, trait == "SLA"))
select = stepwise(reg, criterion = "BIC")
```

```{r, results = "asis"}
tab_model(select)
```

#### logSeedMass
```{r}
reg = lm(formula = mean ~ (dry_sc + rad_sc + Depth_sc + Psand_sc + CEC_sc) : loss_rel + loss_rel - 1, 
          weights = 1/sd, data = subset(df_Delta, trait == "logSeedMass"))
select = stepwise(reg, criterion = "BIC")
```

```{r, results = "asis"}
tab_model(select)
```

#### woodDensity
```{r}
reg = lm(formula = mean ~ (dry_sc + rad_sc + Depth_sc + Psand_sc + CEC_sc) : loss_rel + loss_rel - 1, 
          weights = 1/sd, data = subset(df_Delta, trait == "woodDensity"))
select = stepwise(reg, criterion = "BIC")
```

```{r, results = "asis"}
tab_model(select)
```


## Integrating covariates to the stan model 

```{r}
library(rstanarm)
#https://cran.r-project.org/web/packages/rstanarm/vignettes/lm.html


```