---
title: "What drives the functional response to logging in Amazonia?"
output: 
  html_document:
    theme: yeti
    keep_md: true
    toc: true
    toc_float: true
---

```{r global_options, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning=FALSE, message=FALSE, results='hide',
                      root.dir = "C:/Users/camille.piponiot/gitR/functional_trajectories_TmFO")
library(rstan)
rstan_options(auto_write = TRUE)
options(mc.cores = parallel::detectCores())

library(data.table)
library(ggplot2)
library(gridExtra)
library(tidyr)
library(knitr)
library(ggpubr)

inference_stan = FALSE
```


In this study we model the functional resistance and resilience of trees recruitement to disturbance (here selective logging), and their spatial variation in Amazonia. 

# Data visualisation

## by cohort
```{r, traits_traj, fig.height=10, fig.width=10}
# to get the data: source("C:/Users/camille.piponiot/Google Drive/biodiversity/codes/get_metadata.R")
load("data/sizeData_TmFO.Rdata")
load("data/plotInfo_TmFO.Rdata")
name_traits = c("woodDensity","logSeedMass","SLA","DBH95")

plotInfo = merge(plotInfo, sizeData[,.(tlog = min(year)), .(idplot)], by = "idplot")
plotInfo$tlog[plotInfo$site == "prc"] = 1986
plotInfo$tlog[plotInfo$site == "tor"] = 2006
plotInfo$tlog[plotInfo$site == "cum"] = 2004

### reorganize data -> gather dataframes
sizeData_melt = melt(sizeData, variable.name = "trait", value.name = "value", 
                      measure.vars = grep("D|Seed|SLA",colnames(sizeData)), 
                      variable.factor = FALSE)
dfTraits = sizeData_melt[,.(WMT = weighted.mean(value, agb), agb_tot = sum(agb)),.(idplot, year, trait, size)]

## add plot info (plot size, treatment)
dfTraits = merge(dfTraits, plotInfo, by = c("idplot"))
dfTraits[, agb := agb_tot/idplot.size]

## initial value
md_0 = dfTraits[year <= tlog,.(WMT0 = mean(WMT)),.(site, idplot, trait, treat)]

```

## Traits trajectory {.tabset .tabset-fade .tabset-pills}

Traits chosen: 

- $DBH95$: DBH 95th percentile (per species) as a proxy of the stature

- $logSeedMass$: median seed mass (log-transformed) as a proxy of the dispersal strategy

- $SLA$: median specific leaf area, as a proxy of the assimilation strategy and leaf economic spectrum

- $WD$: median wood density, as a proxy of growth rate and mechanical support

We separate two size classes: 

- Small trees, ie trees $\leq$ 30 cm dbh

- Big trees, ie trees $>$ 50 cm dbh

Here are represented for each size class the mean trait value, weighted by the estimated above ground biomass (AGB) of each individual.

```{r}
dfTraits$size <- factor(dfTraits$size)
levels(dfTraits$size) <- c("small", "big")

graph_traits <- function(sit) {
  ggplot(subset(dfTraits, site == sit), aes(x = year - tlog, y = WMT, 
                                            group = idplot, colour = treat)) + 
    geom_line() + geom_point() + 
    geom_point(data = subset(md_0, site == sit), aes(x = 0, y = WMT0), pch = 1, cex = 5) + 
    ylab("MWT") + xlab("Time since logging (yr)") +
    facet_grid(trait ~ size, scales = "free")
}

tmfo_sites <- as.character(unique(dfTraits$site))
```

### Jenaro 
```{r, fig.height=8, fig.width = 10}
graph_traits("bsl")
```

### Chico Bocao 
```{r, fig.height=10, fig.width = 10}
graph_traits("chb")
```

### Cumaru 
```{r, fig.height=10, fig.width = 10}
graph_traits("cum")
```

### INPA 
```{r, fig.height=10, fig.width = 10}
graph_traits("inp")
```

### Iracema 
```{r, fig.height=10, fig.width = 10}
graph_traits("ira")
```

### La Chonta 
```{r, fig.height=10, fig.width = 10}
graph_traits("lch")
```

### Peteco 
```{r, fig.height=10, fig.width = 10}
graph_traits("pet")
```

### Paracou 
```{r, fig.height=10, fig.width = 10}
graph_traits("prc")
```

### Paragominas 
```{r, fig.height=10, fig.width = 10}
graph_traits("prg")
```

### Tabocal 
```{r, fig.height=10, fig.width = 10}
graph_traits("tbc")
```

### Tapajos 
```{r, fig.height=10, fig.width = 10}
graph_traits("tpj")
```

### Tortue 
```{r, fig.height=10, fig.width = 10}
graph_traits("tor")
```


## Recruits most common genuses and their impact on mean weighted traits {.tabset .tabset-pills}

```{r, fig.height=20, fig.width=12}
load("data/genusData_TmFo.Rdata")
recrData = merge(genusData, genusData[cohort=="recr",.(agbR = sum(agb)),
                                      .(idplot, year, cohort)])

recrData = merge(recrData, plotInfo, by=c("site", "idplot"))
```

```{r}
graph_genus <- function(sit) {
  
  dat = subset(recrData, site == sit)
  ## genuses that at least once represent 10% of the total biomass of recruits
  genus_common = unique(as.character(subset(dat, agb/agbR > 0.1 & !(genus %in% c("Oenocarpus", "Indet", NA)))$genus))
  dat$genus2 = as.character(dat$genus)
  dat$genus2[!(dat$genus %in% genus_common)] <- "other"
  dat$genus2 = factor(dat$genus2, levels = c(as.character(genus_common), "other") )
  
  dat_agb = dat[, .(agb = sum(agb)), .(genus2,idplot,year,agbR)]
  dat = merge(dat_agb, dat[, .(woodDensity = weighted.mean(woodDensity, agb),
                                    logSeedMass = weighted.mean(logSeedMass, agb),
                                    SLA = weighted.mean(SLA, agb)), .(genus2)])
  
  g1 = ggplot(dat, aes(x=year, y=agb, fill=genus2)) + 
    geom_area(colour = 1) + facet_wrap( ~ idplot)
  g1
  ## mean traits (weighted by biomass) of recruited genuses in control plots -> reference (= white color in plots)
  # ref = dat[treat=="ctrl", .(woodDensity = weighted.mean(woodDensity, agb),
  #                            logSeedMass = weighted.mean(logSeedMass, agb),
  #                            SLA = weighted.mean(SLA, agb))]
}

```

genus2: genuses with > 10% of the recruits biomass

### Jenaro 
```{r, fig.height=4, fig.width = 10}
graph_genus("bsl")
```

### Chico Bocao 
```{r, fig.height=4, fig.width = 6}
graph_genus("chb")
```

### Cumaru 
```{r, fig.height=10, fig.width = 10}
graph_genus("cum")
```

### INPA 
```{r, fig.height=10, fig.width = 10}
graph_genus("inp")
```

### Iracema 
```{r, fig.height=10, fig.width = 10}
graph_genus("ira")
```

### La Chonta 
```{r, fig.height=10, fig.width = 10}
graph_genus("lch")
```

### Peteco 
```{r, fig.height=10, fig.width = 10}
graph_genus("pet")
```

### Paracou 
```{r, fig.height=10, fig.width = 10}
graph_genus("prc")
```

### Paragominas 
```{r, fig.height=10, fig.width = 10}
graph_genus("prg")
```

### Tabocal 
```{r, fig.height=10, fig.width = 10}
graph_genus("tbc")
```

### Tapajos 
```{r, fig.height=10, fig.width = 10}
graph_genus("tpj")
```

### Tortue 
```{r, fig.height=10, fig.width = 10}
graph_genus("tor")
```


## Calibrating the stan model

For each trait $k$, the mean biomass-weighted trait of small trees (MWT) trajectory at census $c$ in plot $p$ in site $s$ was modelled as:

$$  T_{k,c,p,s} = \underbrace{T0_{k,p,s}}_{\text{pre-logging value}}\left(1  +   \underbrace{\left(\Delta_{k,p,s} \cdot  \frac{t_c}{tmax_{k,p,s}} \cdot exp\left(1-\frac{t_c}{tmax_{k,p,s}}\right)\right)^{\theta_{k,p,s}}}_{\text{boom and bust pattern}} \right)$$

with 

- $c$ the census, and $t_c$ the time since logging (in years) at census $c$, 

- $T0_{k,p,s}$ the pre-logging trait value of the community in plot $p$ in site $s$.

- $\Delta_{k,p,s}$ is the maximum value (relative to the final value) of the hump that follows logging and the creation of logging gaps. $\Delta_{k,p,s} \sim \mathcal{N} (\mu_{\Delta ks}, \sigma_{\Delta}^2)$, with $\mu_{\Delta ks}$ and $\sigma_{\Delta}$ are respectively the mean and standard deviation of $\Delta$ for trait $k$ in site $s$.

- $tmax_{k,p,s}$ is the time when the maximum value of the hump is reached. $tmax_{k,p} = 0$ for control plots and in logged plots $tmax_{k,p} \sim \mathcal{N} (\mu m_{ k}, \sigma m^2)$ where $\mu m_{k}$ and $\sigma m$ are respectively the mean and standard deviation of $tmax$ for trait $k$.

- $\theta_k$ is a shape parameter that controls the width of the hump; when it increases, the hump is narrower.  

```{r}
if (inference_stan){
  
  df = subset(dfTraits, size=="small")
  name_traits = unique(df$trait)
  df = dcast(df, site + idplot + tlog + year + idplot.size ~ trait , value.var = "WMT")
  df0 = subset(df, year <= tlog)
  df = subset(df, year > tlog)
  
  N = nrow(df)
  K = length(name_traits)
  np = as.numeric(as.factor(df$idplot))
  P = length(unique(np))
  ns = as.numeric(as.factor(df$site))
  S = length(unique(ns))
  ps = unique(cbind(ns, np))[,1]
  plot_size = unique(cbind(df$idplot.size, np))[,1]
  
  t = df$year - df$tlog
  TR = as.matrix(df[, name_traits, with = FALSE])
  dfT0 = melt(df0, measure.vars = name_traits, id.vars = "idplot")
  dfT0 = dfT0[,.(value = mean(value)),.(variable,idplot)]
  T0 = as.matrix(dcast(dfT0, idplot ~variable, value.var = "value")[,-"idplot"])
  
  stan("codes/recruits_model.stan")
  
}


```