---
title: "Testing all covariates effect on the functional resistance"
output: 
  html_document:
    theme: yeti
    keep_md: true
    toc: true
    toc_float: true
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning=FALSE, message=FALSE, results='hide',
                      root.dir = "C:/Users/camille.piponiot/gitR/functional_trajectories_TmFO")
library(rstan)
rstan_options(auto_write = TRUE)
options(mc.cores = parallel::detectCores())
inference_stan = FALSE
with_initial_traits = FALSE ## inference does not include initial trait value (changed later)

library(data.table)
library(ggplot2)
library(gridExtra)
library(knitr)
library(ggpubr)
library(factoextra)
library(GGally)
```

```{r load_data}
## study sites
load("data/site_coord.Rdata")
load("data/plotInfo_TmFO.Rdata")
site_coord = subset(site_coord, site %in% unique(plotInfo$site))
rownames(site_coord) = site_coord$site

## trait data
load("data/sizeData_TmFO.Rdata")
name_traits = c("woodDensity","logSeedMass","SLA","DBH95")

plotInfo = merge(plotInfo, sizeData[,.(tlog = min(year)), .(idplot)], by = "idplot")
plotInfo$tlog[plotInfo$site == "prc"] = 1986
plotInfo$tlog[plotInfo$site == "tor"] = 2006
plotInfo$tlog[plotInfo$site == "cum"] = 2004

### reorganize data -> gather dataframes
sizeData_melt = melt(sizeData, variable.name = "trait_FD", value.name = "value", 
                     measure.vars = grep("DBH95|wood|Seed|SLA",colnames(sizeData)), 
                     variable.factor = FALSE)
sizeData_melt$trait = gsub("FD","",sizeData_melt$trait_FD)
sizeData_melt$measure = c("WMT","FD")[grepl("FD",sizeData_melt$trait_FD)+1]
sizeData_melt = dcast(sizeData_melt, idplot + site + year + size + agb + 
                        FDtot + trait ~ measure, value.var = "value")

dfTraits = sizeData_melt[,.(WMT = weighted.mean(WMT, agb), agb_tot = sum(agb), 
                            FDtot = mean(FDtot), FD = mean(FD)),.(idplot, year, trait, size)] ## FD must be calculated for each new plot

## add plot info (plot size, treatment)
dfTraits = merge(dfTraits, plotInfo, by = c("idplot"))
dfTraits[, agb := agb_tot/idplot.size]
# size name
dfTraits$size <- factor(dfTraits$size)
levels(dfTraits$size) <- c("small", "big")

## df_covar
load("data/parsSmall.Rdata")
pars_melt = melt(pars_small, measure.vars = c("Delta","tmax","theta"), 
                 id.vars = c("iter","site","trait","idplot"), variable.name = "parameter")
df_covar = merge(pars_melt[,.(mean = mean(value), sd = sd(value)), .(site, trait, parameter, idplot)],
                 site_coord, by = "site")
## agb loss
df_agb = dfTraits[trait == "SLA" & treat != "ctrl",
                  .(agb = sum(agb_tot)), .(site,idplot,year,tlog)]
df_dist = df_agb[,.(agb0 = mean(agb[year <= tlog]), 
                    agbmin = min(agb[year>tlog & year <= tlog + 4])),.(idplot)]
df_dist[, loss_rel := (agb0-agbmin)/agb0]
df_dist[, loss_abs := (agb0-agbmin)]
df_covar = merge(df_covar,  df_dist, by = "idplot")

df_covar = melt(df_covar, id.vars = c("site", "idplot","trait","parameter","mean","sd"), 
                measure.vars = c(colnames(site_coord)[6:length(site_coord)],"loss_rel","loss_abs"), 
                variable.name = "covariate")

## df_Delta
df_Delta = subset(df_covar, parameter == "Delta")
df_Delta = dcast(unique(df_Delta[,c("site","idplot","mean","sd","covariate","value","trait")]), 
                 site + idplot + trait + mean + sd ~ covariate, value.var = "value")
## scale
cov_sc = data.table(scale(site_coord[,6:ncol(site_coord)]))
colnames(cov_sc) = paste0(colnames(cov_sc), "_sc")
cov_sc$site = rownames(site_coord)
df_Delta = merge(df_Delta, cov_sc, by = "site")
```

```{r}
# select censuses previous to logging
md0 = subset(dfTraits, year <= tlog & size == "small")
md0 = md0[,.(FD = mean(FD), FDtot = mean(FDtot), WMT = mean(WMT)), .(site,idplot,trait)]

# scale covariates
md0 = merge(md0, md0[, .(FD_sc = (FD-mean(FD))/sd(FD), 
                         FDtot_sc = (FDtot-mean(FDtot))/sd(FDtot),
                         WMT_sc = (WMT-mean(WMT))/sd(WMT), 
                         idplot, site), 
                     .(trait)],
            by = c("idplot","site","trait"))

if (length(df_Delta$WMT)==0) 
  df_Delta = merge(df_Delta, md0, by = c("site","idplot","trait"))
site_coord_trait = merge(data.table(site_coord), df_Delta[, .(FD = mean(FD), 
                                                              FDtot = mean(FDtot),
                                                              WMT = mean(WMT)), 
                                                          .(site, trait)], by = "site")
site_coord = merge(site_coord, dcast(site_coord_trait, site ~ trait, value.var = "WMT"), by = "site")
dfFD = dcast(site_coord_trait, site ~ trait, value.var = "FD")
colnames(dfFD) = c("site",paste0("FD", c("dbh","sla","seed","wd")))
site_coord = merge(site_coord, dfFD, by = "site")
```

# Correlations 

```{r}
## change cec value in Paracou: too high
## find better estimate later?
site_coord$CEC[site_coord$site=="prc"] = site_coord$CEC[site_coord$site=="tor"]
```

```{r, fig.height=10, fig.width=10}
ggcorr(site_coord[,6:ncol(site_coord)], nbreaks = 5, label = TRUE, high="red",low="red",middle = "blue") + 
  labs(title = "Covariates correlation") 
```

# Effect of covariates (pairwise)

```{r}
all_covariates = gsub("_sc", "", grep("_sc", colnames(df_Delta), value = T))
all_covariates = all_covariates[-grep(c("WMT|FD"), all_covariates)]

if (inference_stan){
  for(cov_names in all_covariates) {
    source("codes/inference_covariates_model.R")
    save(pars_cov, file = paste0("data/temp/parsCov_", cov_names,".Rdata"))
  }
}

df_lambda = data.table(expand.grid(iter = 1:1000, trait = c("DBH95","SLA","logSeedMass","woodDensity")))

for(cov_names in all_covariates) {
  load(paste0("data/temp/parsCov_", cov_names,".Rdata"))
  df_lambda = merge(df_lambda, 
                    unique(pars_cov[,c("iter","trait",grep("lambda_", colnames(pars_cov), value=T)), with = FALSE]), 
                    by = c("iter","trait"))
}
colnames(df_lambda) = gsub(pattern = "lambda_", replacement = "", x = colnames(df_lambda))

df_lambda = melt(df_lambda, id.vars = c("iter","trait"), value.name = "lambda", variable.name = "covar")

df_lambda = merge(df_lambda, 
                  df_lambda[,.(sign = (quantile(lambda, 0.05) > 0 | 
                                         quantile(lambda, 0.95) < 0)),
                            .(trait,covar)], by = c("trait","covar"))
```

```{r, fig.height=10, fig.width = 10}
ggplot(df_lambda, aes(x=trait, fill=sign, y = lambda)) + 
  geom_hline(yintercept = 0, lty=2) + 
  geom_violin() + 
  facet_wrap(~ covar, ncol = 4)
```

# By pairs {.tabset .tabset-pills}

smort and Psand are the only good predictor of the WD resistance.
We test the effect of smort with seas or CEC (not too correlated)
[ maybe add in supp mat the results with WMT -> not only due to initial trait value]

## smort + agb productivity 

```{r}
cov_names = c("agbp","smort")

if (inference_stan){
  source("codes/inference_covariates_model.R")
  save(pars_cov, file = "data/parsCov_agbp_smort.Rdata")
} else { load("data/parsCov_agbp_smort.Rdata")}

```

```{r}
df = unique(pars_cov[, c("trait", grep("lambda", colnames(pars_cov), value = TRUE)), with = F])
source("codes/plot_lambda.R")
df_covar$covar_typ = 1
plot_lambda(df$trait, df[,-"trait"], unique(df_covar[,c("covariate","covar_typ")])) + 
  labs(title = "lambdas") + theme(legend.position="none")
```

## smort + seas 

```{r}
cov_names = c("seas","smort")

if (inference_stan){
  source("codes/inference_covariates_model.R")
  save(pars_cov, file = "data/parsCov_seas_smort.Rdata")
} else { load("data/parsCov_seas_smort.Rdata")}

```

```{r}
df = unique(pars_cov[, c("trait", grep("lambda", colnames(pars_cov), value = TRUE)), with = F])
source("codes/plot_lambda.R")
df_covar$covar_typ = 1
plot_lambda(df$trait, df[,-"trait"], unique(df_covar[,c("covariate","covar_typ")])) + 
  labs(title = "lambdas") + theme(legend.position="none")
```

## smort + CEC

```{r}
cov_names = c("CEC","smort")

if (inference_stan){
  source("codes/inference_covariates_model.R")
  save(pars_cov, file = "data/parsCov_CEC_smort.Rdata")
} else { load("data/parsCov_CEC_smort.Rdata") }

```

```{r}
df = unique(pars_cov[, c("trait", grep("lambda", colnames(pars_cov), value = TRUE)), with = F])
source("codes/plot_lambda.R")
df_covar$covar_typ = 1
plot_lambda(df$trait, df[,-"trait"], unique(df_covar[,c("covariate","covar_typ")])) + 
  labs(title = "lambdas") + theme(legend.position="none")
```

## smort + CWD

```{r}
cov_names = c("cwd","smort")

if (inference_stan){
  source("codes/inference_covariates_model.R")
  save(pars_cov, file = "data/parsCov_cwd_smort.Rdata")
} else { load("data/parsCov_cwd_smort.Rdata")}

```

```{r}
df = unique(pars_cov[, c("trait", grep("lambda", colnames(pars_cov), value = TRUE)), with = F])
source("codes/plot_lambda.R")
df_covar$covar_typ = 1
plot_lambda(df$trait, df[,-"trait"], unique(df_covar[,c("covariate","covar_typ")])) + 
  labs(title = "lambdas") + theme(legend.position="none")
```

## smort + rad

```{r}
cov_names = c("rad","smort")

if (inference_stan){
  source("codes/inference_covariates_model.R")
  save(pars_cov, file = "data/parsCov_rad_smort.Rdata")
} else { load("data/parsCov_rad_smort.Rdata")}

```

```{r}
df = unique(pars_cov[, c("trait", grep("lambda", colnames(pars_cov), value = TRUE)), with = F])
source("codes/plot_lambda.R")
df_covar$covar_typ = 1
plot_lambda(df$trait, df[,-"trait"], unique(df_covar[,c("covariate","covar_typ")])) + 
  labs(title = "lambdas") + theme(legend.position="none")
```

# with initial trait value

## with initial trait value only (no other covariate)

```{r}
if (inference_stan) {
  model_covar = stan("codes/model_traits.stan")
  
  pars_cov = data.table(iter = 1:1000, do.call(cbind, rstan::extract(model_covar, 
                                                                     pars = c("sigma_Delta", "sigma_tmax", 
                                                                              "mu_theta", "sigma_theta", "lp__" )))[1:1000])
  pars_covb = data.table(Delta = c(rstan::extract(model_covar, pars= "Delta")[[1]][1:1000,,]), 
                         theta = c(rstan::extract(model_covar, pars= "theta")[[1]][1:1000,,]),
                         tmax = c(rstan::extract(model_covar, pars= "tmax")[[1]][1:1000,,]), 
                         iter = rep(1:1000, P*K), 
                         p = rep(rep(1:P, each = 1000), K), 
                         s = rep(rep(ps, each = 1000), K), 
                         k = rep(1:K, each = 1000*P))
  pars_cov = merge(pars_cov, pars_covb, by = "iter")
  pars_covb = data.table(sigmaT = c(rstan::extract(model_covar, pars= "sigmaT")[[1]][1:1000,]),
                         sigma_Delta = c(rstan::extract(model_covar, pars= "sigma_Delta")[[1]][1:1000,]),
                         lambda0 = c(rstan::extract(model_covar, pars= "lambda0")[[1]][1:1000,]),
                         iter = rep(1:1000, K), 
                         k = rep(1:K, each = 1000))
  pars_covb$lambda_WMT = c(rstan::extract(model_covar, pars= "gamma")[[1]][1:1000,])
  
  
  pars_cov = merge(pars_cov, pars_covb, by = c("iter", "k"))
  pars_covb = data.table(mu_tmax = c(rstan::extract(model_covar, pars= "mu_tmax")[[1]][1:1000,,]), 
                         iter = rep(1:1000, S*K), 
                         s = rep(rep(1:S, each = 1000), K), 
                         k = rep(1:K, each = 1000*S))
  pars_cov = merge(pars_cov, pars_covb, by = c("iter","s","k"))
  pars_cov$trait = name_traits[pars_cov$k]
  pars_cov$site = levels(as.factor(as.character(df$site)))[pars_cov$s]
  pars_cov$idplot = levels(as.factor(as.character(df$idplot)))[pars_cov$p]
  save(pars_cov, file = "data/parsCov_traitsOnly.Rdata")
} else {load("data/parsCov_traitsOnly.Rdata")}
```

```{r}
df = unique(pars_cov[, c("trait", grep("lambda", colnames(pars_cov), value = TRUE)), with = F])
source("codes/plot_lambda.R")
df_covar$covar_typ = 1
plot_lambda(df$trait, df[,-"trait"], unique(df_covar[,c("covariate","covar_typ")])) + 
  labs(title = "lambdas") + theme(legend.position="none")
```


## with trait value and covariates

```{r}
with_initial_traits = TRUE

with_initial_FD = FALSE
cov_names = c("seas","smort")

if (inference_stan){
  source("codes/inference_covariates_model.R")
  save(pars_cov, file = "data/parsCov_seas_smort_traits.Rdata")
} else { load("data/parsCov_seas_smort_traits.Rdata")}

```

```{r}
df = unique(pars_cov[, c("trait", grep("lambda", colnames(pars_cov), value = TRUE)), with = F])
source("codes/plot_lambda.R")
df_covar$covar_typ = 1
plot_lambda(df$trait, df[,-"trait"], unique(df_covar[,c("covariate","covar_typ")])) + 
  labs(title = "lambdas") + theme(legend.position="none")
```

> After checking that adding the initial trait value does not modify previous results, I choose to remove them. The 2 reasons are: (i) I do not want to comment on their effect; (ii) their effect may be poorly estimated, as the value is different when we add the covariates or not. The only trait that has an effect when only intial WMT are considered is the DBH95, but this effect disappears when covariates are added, so it could be an effect of intersite differences explained by environamental gradients, rather than the trait itself.  

# with initial functional diversity

## with initial functional diversity only (no other covariate)

```{r}

with_initial_FD = TRUE

if (inference_stan) {
  
  ## change initial trait value by initial functional diversity, scaled
  df_FD = df_Delta[,c("site","idplot","trait","FD_sc")]
  df_FD = dcast(df_FD, site + idplot ~ trait, value.var = "FD_sc")[,-c("site","idplot")]
  T0_sc = as.matrix(scale(df_FD))
  model_covar = stan("codes/model_traits.stan")
  
  pars_cov = data.table(iter = 1:1000, do.call(cbind, rstan::extract(model_covar, 
                                                                     pars = c("sigma_Delta", "sigma_tmax", 
                                                                              "mu_theta", "sigma_theta", "lp__" )))[1:1000])
  pars_covb = data.table(Delta = c(rstan::extract(model_covar, pars= "Delta")[[1]][1:1000,,]), 
                         theta = c(rstan::extract(model_covar, pars= "theta")[[1]][1:1000,,]),
                         tmax = c(rstan::extract(model_covar, pars= "tmax")[[1]][1:1000,,]), 
                         iter = rep(1:1000, P*K), 
                         p = rep(rep(1:P, each = 1000), K), 
                         s = rep(rep(ps, each = 1000), K), 
                         k = rep(1:K, each = 1000*P))
  pars_cov = merge(pars_cov, pars_covb, by = "iter")
  pars_covb = data.table(sigmaT = c(rstan::extract(model_covar, pars= "sigmaT")[[1]][1:1000,]),
                         sigma_Delta = c(rstan::extract(model_covar, pars= "sigma_Delta")[[1]][1:1000,]),
                         lambda0 = c(rstan::extract(model_covar, pars= "lambda0")[[1]][1:1000,]),
                         iter = rep(1:1000, K), 
                         k = rep(1:K, each = 1000))
  pars_covb$lambda_FD = c(rstan::extract(model_covar, pars= "gamma")[[1]][1:1000,])
  
  
  pars_cov = merge(pars_cov, pars_covb, by = c("iter", "k"))
  pars_covb = data.table(mu_tmax = c(rstan::extract(model_covar, pars= "mu_tmax")[[1]][1:1000,,]), 
                         iter = rep(1:1000, S*K), 
                         s = rep(rep(1:S, each = 1000), K), 
                         k = rep(1:K, each = 1000*S))
  pars_cov = merge(pars_cov, pars_covb, by = c("iter","s","k"))
  pars_cov$trait = name_traits[pars_cov$k]
  pars_cov$site = levels(as.factor(as.character(df$site)))[pars_cov$s]
  pars_cov$idplot = levels(as.factor(as.character(df$idplot)))[pars_cov$p]
  save(pars_cov, file = "data/parsCov_funcDivOnly.Rdata")
} else {load("data/parsCov_funcDivOnly.Rdata")}
```

```{r}
df = unique(pars_cov[, c("trait", grep("lambda", colnames(pars_cov), value = TRUE)), with = F])
source("codes/plot_lambda.R")
df_covar$covar_typ = 1
plot_lambda(df$trait, df[,-"trait"], unique(df_covar[,c("covariate","covar_typ")])) + 
  labs(title = "lambdas") + theme(legend.position="none")
```

## with initial fdiv and trait and covariates

```{r}
with_initial_traits = TRUE
with_initial_FD = TRUE
cov_names = c("seas","smort")

if (inference_stan){
  source("codes/inference_covariates_model.R")
  save(pars_cov, file = "data/parsCov_seas_smort_traits_FD.Rdata")
} else { load("data/parsCov_seas_smort_traits_FD.Rdata")}

```

```{r}
df = unique(pars_cov[, c("trait", grep("lambda", colnames(pars_cov), value = TRUE)), with = F])
source("codes/plot_lambda.R")
df_covar$covar_typ = 1
plot_lambda(df$trait, df[,-"trait"], unique(df_covar[,c("covariate","covar_typ")])) + 
  labs(title = "lambdas") + theme(legend.position="none")
```

# Cross validation

## seas + smort 

```{r}
with_initial_traits = FALSE
cov_names = c("seas","smort")

if (inference_stan){
  source("codes/inference_covariates_model.R")
} 

load("data/parsCov_seas_smort_traits.Rdata")
pars_loo = data.table(site = "all",
                      unique(pars_cov[,c("iter", "k","lambda0", "lambda_seas", 
                                         "lambda_smort", "trait")]))

for (sit in unique(df_covar$site)) {
  load(paste0("data/pars_cov_without_", sit, ".Rdata"))
  pars_loo = rbind(pars_loo, data.table(site = sit, pars_cov))
} 

pars_loo = melt(pars_loo, variable.name = "covar", value.name = "lambda",
                measure.vars = grep("lambda", colnames(pars_loo), value = TRUE))
pars_loo$covar = gsub("lambda|_", "", pars_loo$covar)
pars_loo$covar[pars_loo$covar == "0"] = "intercept"

ci_loo = pars_loo[, .(inf = quantile(lambda, 0.025), 
                      med = quantile(lambda, 0.5), 
                      sup = quantile(lambda, 0.975)), 
                  .(covar, site, trait)]
```

```{r, fig.height = 10, fig.width = 8}
ggplot(ci_loo, aes(x = site, y = med, ymin = inf, ymax = sup)) + 
  geom_hline(yintercept = 0, lty = 2) +
  coord_flip() + 
  geom_pointrange() +
  facet_grid( trait ~ covar)
```

> Paragominas has some strange results, and I do not have a high confidence in raw data from this site (problem with small-tree subplot design): maybe we should remove it. 

> La Chonta seems to be a very unique site and has a large impact on final results (removing it results in large differences in parameter estimation). 

> Alternative idea: fit the model with all sites, but one site without covariates, and test how well the Delta is estimated with new lambda values